Figure 3a

### 2018-05-01 inst/lncTAM34a_KD/lncTAM34a_KD.Rmd

We hypothesized that miR34a asRNA may function to mediate the levels of its sense gene, miR34a. To this end, we stably over-expressed miR34a asRNA, using lentiviral transduction, in 3 different p53-null cell lines and monitored the effects on miR34a.

@@

_miR34a asRNA cloning_

RNA was harvested from U2OS cells using the RNeasy mini kit (Qiagen) and subsequently treated with DNase (Ambion Turbo DNA-free, Life Technologies). 500ng RNA was used for cDNA synthesis using MuMLV (Life Technologies) and a 1:1 mix of oligo(dT) and random nanomers. PCR was run using the miR34aAS_cloning_F4 and miR34aAS_cloning_Ex3_1 primers for 40 cycles at annealing temperature 58˚C and run on an agarose gel. Bands were excised and extracted using the QIAquick gel extraction kit (Qiagen). PCR products were cloned using the Strata Clone kit (Aligent) and sequenced (results below). The insert for the lentiviral construct (p-HIV-I2AG-U6) was subsequently sub-cloned from the sequenced plasmid using primers with PacI and NheI cleavage sites. The cloned transcript largley mirrors the GENCODE annotated transcript.

@@

_Cloning primers_

@@

primers <- data.frame(
    name=c(
        "miR34aAS_cloning_F4",
        "miR34aAS_cloning_Ex3_1"
    ),
    sequence=c(
        "ACGCGTCTCTCCAGCCCGGGAT",
        "AATGATGGCCGCAACTAATGACGG"
    )
)
primers

@@

RNA was extracted using the RNeasy mini kit (Qiagen) and subsequently treated with DNase (Ambion Turbo DNA-free, Life Technologies). 500ng RNA was used for cDNA synthesis using MuMLV (Life Technologies) and a 1:1 mix of oligo(dT) and random nanomers. QPCR was carried out using KAPA 2G SYBRGreen (Kapa Biosystems) using the Applied Biosystems 7900HT machine with the cycling conditions: 95 °C for 3 min, 95 °C for 3 s, 60 °C for 30 s. QPCR for miRNA expression analysis was performed according to the protocol for the TaqMan microRNA Assay kit (Life Technologies) with primer/probe sets purchased from the same company (TaqMan® MicroRNA Assay, hsa-miR-34a, human and Control miRNA Assay, RNU48, human) and the same cycling scheme as above. QPCR primers are shown below.

@@

Two experimental (technical) replicates were included in each QPCR run and delta ct was calculated for each sample using the mean of the gene of interest's technical replicates and the house keeping gene's technical replicates. delta-delta ct was calculated for each sample by subtracting the median dct value for the corresponding mock samples. Five independant experiments were performed in total. The Student's t-test was used to compare the mock vs miR34a asRNA overexpressing group's delta-delta ct values for both genes.

_QPCR primers_

primers <- data.frame(
    name=c(
        "ß-actin Fwd",
        "ß-actin Rev",
        "miR34a_asF1",
        "miR34a_asR1"

    ),
    sequence=c(
        "AGGTCATCACCATTGGCAATGAG",
        "CTTTCGGGATGTCCACGTCA",
        "AGCGGCATCTCCTCCACCTGAAA",
        "TTGCCTCGTGAGTCCAAGGAGAAT"

    )
)
primers

@@

#data <- getData("Figure 3a")
data <- read_tsv('~/Github/miR34a_asRNA_project/data-raw/lncTAM34a_KD.txt')

@@

data <- data %>%
  technicalMeans(., groups) %>%
  dct(., gois, Actin, groups, F)

@@

data <- data %>%
  ddct(., logical, groups, F) %>%
  folds(.)

@@

#calculate 95% CI and mean on linear scale
stats <- data %>%
  filter(siRNA == "si-lncTAM34a") %>%
  group_by(GOI, siRNA) %>%
  summarize(
    CI95l.linear = t.test(fold)$conf.int[1], 
    CI95h.linear = t.test(fold)$conf.int[2]
  ) %>%
  ungroup() %>%
  full_join(stats, by = c("siRNA", "GOI"))

@@

#vertical lines and labels in plot
vl <- stats %>%
  mutate(
    x = case_when(
      GOI == "miR34a HG" ~ 2 - 0.25,
      GOI == "lncTAM34a" ~ 1 - 0.25
    ),
    xend = case_when(
      GOI ==  "miR34a HG" ~ 2 + 0.25,
      GOI == "lncTAM34a" ~ 1 + 0.25
    ),
    label = if_else(GOI == "miR34a HG", 2, 1), 
    max = 1
  )
```

@@

# path <- file.path("~/GitHub/miR34a_asRNA_project/inst", fileMap(type = "pdf")["Figure 3a"][[1]])
# ggsave(
#    plot = figure,
#    filename = path,
#    device = cairo_pdf,
#    height = 60,
#    width = 83,
#    units = "mm"
# )
```

<div class="caption">
miR34a asRNA and miR34a levels in PC3, Skov3, and Saos2 cell lines stably over-expressing miR34a asRNA compared to mock over-expressing control. The points represent the values obtained from each independant experiment (n >= 3). 95% confidence interval (vertical black lines), mean (horizontal black lines), and p-values are shown with the horizontal line under the p-value indicating the comparison that was tested.
</div>

@@

miR34a asRNA over-expression causes a concominant up-regulation of miR34a in all 3 cell lines. This indicates that miR34a asRNA concordantly regulates miR34a.

>>>>

We hypothesized that miR34a asRNA may function to mediate the levels of its sense gene, miR34a. To this end, we utilized
siRNA targeting exon 2 of the miR34a asRNA transcript to lower its expression levels and observe possible effects on 
miR34a HG in U2OS cells.

@@

_Cell culture, transfection and luminescence quantification_

All cell lines were cultured at 5% CO2 and 37° C with U2OS cells cultured in McCoy’s 5a (Life Technologies). Growth mediums were supplemented with 10% heat-inactivated FBS and 50 μg/ml of streptomycin and 50 μg/ml of penicillin. Cells were plated at 25,000 cells per well in a 12-well plate and cultured overnight. The following day cells were transfected with 10nM si-Ctl or si-lncTAM34a using the standard lipofektamine 2000 (Life Technologies) protocol. si-Ctl was purchased from Qiagen (cat. nr. 1027310) and si-lncTAM34a was purchased from Eurofins. The sequence for si-Ctl is not available from the manufacturer although the sequence for si-lncTAM34a is shown below. Cells were harvested for RNA extraction 48 hours after transfection.

@@

_siRNA_

@@

siRNA <- data.frame(
  name = c("si-lncTAM34a"),
  sequence = c("GGG AGA AGA CGA UUC UUU")
)
siRNA

@@

RNA was extracted using the RNeasy mini kit (Qiagen) and subsequently treated with DNase (Ambion Turbo DNA-free, Life Technologies). 500ng RNA was used for cDNA synthesis using MuMLV (Life Technologies) and a 1:1 mix of oligo(dT) and random nanomers. QPCR was carried out using KAPA 2G SYBRGreen (Kapa Biosystems) using the Applied Biosystems 7900HT machine with the cycling conditions: 95 °C for 3 min, 95 °C for 3 s, 60 °C for 30 s. QPCR primers are shown below.

_QPCR Primers_

primers <- data.frame(
    name=c(
        "miR34a asRNA F1",
        "miR34a asRNA R1",
        "miR34aHG_F",
        "miR34aHG_R",
        "B-actin_F",
        "B-actin_R"
    ),
    sequence=c(
        "AGC GGC ATC TCC TCC ACC TGA AA",
        "TTG CCT CGT GAG TCC AAG GAG AAT",
        "TCT GCT CCA GTG GCT GAT GAG AAA",
        "GTT CAC TGG CCT CAA AGT TGG CAT",
        "AGG TCA TCA CCA TTG GCA ATG AG",
        "CTT TGC GGA TGT CCA CGT CA"
    )
)
primers

@@

Three independant experiments (biological replicates) were performed in total measuring miR34a HG whereas miR34a asRNA was measured in 4 biological replicates. The Student's t-test was used to compare the si-Ctl and si-lncTAM34a conditions log2 fold change for both genes.

@@

data <- getData("Figure 2c")
#data <- read_tsv('~/Github/miR34a_asRNA_project/data-raw/lncTAM34a_KD.txt')

@@

data <- data %>%
  gather(Ct, ctMean, -(gene:`Biological Replicate`)) %>%
  group_by_at(groups) %>%
  mutate(`Technical Replicate` = 1:n()) %>%
  dct(., gois, Actin, c(groups, 'Technical Replicate'), F)

@@

data <- data %>%
  mutate(br = `Biological Replicate`) %>%
  ddct(., logical, c(groups, "br")) %>%
  filter(`Technical Replicate` == 1) %>%
  select(-`Technical Replicate`) %>%
  distinct() %>%
  na.omit() %>%
  folds(.)

@@

#calculate 95% CI and mean on linear scale
stats <- data %>%
  #filter(siRNA == "si-lncTAM34a") %>%
  group_by(GOI, siRNA) %>%
  summarize(
    CI95l.linear = t.test(fold)$conf.int[1], 
    CI95h.linear = t.test(fold)$conf.int[2]
  ) %>%
  ungroup() %>%
  full_join(stats, by = c("siRNA", "GOI"))

@@

#vertical lines and labels in plot
vl <- stats %>%
  mutate(
    x = case_when(
      GOI == "miR34a HG" ~ 2 - 0.25,
      GOI == "lncTAM34a" ~ 1 - 0.25
    ),
    xend = case_when(
      GOI ==  "miR34a HG" ~ 2 + 0.25,
      GOI == "lncTAM34a" ~ 1 + 0.25
    ),
    label = if_else(GOI == "miR34a HG", 2, 1), 
    max = max(CI95h.linear)
  )
```

@@

path <- file.path("~/GitHub/miR34a_asRNA_project/inst", fileMap(type = "pdf")["Figure 2c"][[1]])
ggsave(
  plot = figure,
  filename = path,
  device = cairo_pdf,
  height = 60,
  width = 83,
  units = "mm"
)
```

<div class="caption">
lncTAM34a (n = 4) and miR34a HG (n = 3) levels after 48 hour siRNA-mediated knock down of lncTAM34a in U2OS cells.
</div>

@@

miR34a asRNA KD gives rise to a ~20% decrease in miR34a HG expression levels.
